Reproducible Research Assignment 1
=============================================

```{r initial setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
Loading and preprocessing the data
```{r}
urlzip <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip"
download.file(urlzip, destfile = "./activitymonitoringzip")
unzip("activity_data.zip", exdir = "./activitymonitoring")
activity<-read.csv("activity.csv",header=TRUE,na.strings="NA")
head(activity)
str(activity)
```
Convert date to date format
```{r}
activity$date<-as.Date(activity$date)
```
Histogram of the total number of steps taken per day
```{r}
library(ggplot2)
aggpd <- aggregate(steps ~ date, FUN = sum, data = activity)
hist(aggpd$steps)
```









Mean number of steps taken per day 
```{r}
aggmean <- mean(aggpd$steps)
aggmean
```
Median number of steps taken per day
```{r}
aggmedian <- median(aggpd$steps)
aggmedian
```
Daily average activity pattern
```{r}
fiveminuteav <- aggregate(steps~interval, data=activity, FUN=mean, na.rm=TRUE)
plot(x = fiveminuteav$interval, y = fiveminuteav$steps, type = "l") 
```










Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps
```{r}
max_steps <- max(fiveminuteav$steps)
for (i in 1:288) 
{
    if (fiveminuteav$steps[i] == max_steps)
        five_minute_interval_at_max_steps <- fiveminuteav$interval[i]
}
five_minute_interval_at_max_steps
```
Total number of missing values
```{r}
sum(is.na(activity))
```
Missing value imputation strategy
```{r}
impute.activity <- activity
impute.activity$steps[is.na(impute.activity$steps)] <- mean(impute.activity$steps,na.rm=TRUE)
impute.activity$steps <- as.numeric(impute.activity$steps)
impute.activity$interval <- as.numeric(impute.activity$interval)
```
New dataset that is equal to the original dataset but with the missing data filled in
```{r}
summary(impute.activity)
```
Make a histogram of the total number of steps taken each day
```{r}
aggpdnoNA <- aggregate(steps ~ date, FUN = sum, data = impute.activity)
hist(aggpdnoNA$steps)
```








Mean total number of steps taken per day (after impute)
```{r}
aggmeannoNA<-mean(aggpdnoNA$steps)
aggmeannoNA
```
Median total number of steps taken per day (after impute)
```{r}
aggmediannoNA<-median(aggpdnoNA$steps)
aggmediannoNA
```
Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day
```{r}
impute.activity$day <- ifelse(weekdays(impute.activity$date) %in% c("Saturday","Sunday"), "weekday", "weekend")
```
Panel plot time series by day
```{r}
library(dplyr)
impute.day <- impute.activity %>% group_by(interval,day) %>% summarise(mean.step=mean(steps))
weekdayinterval <- ggplot(impute.day, aes(x=interval, y=mean.step, color=day)) + 
  facet_grid(day~.) +
  geom_line() + 
  labs(title="Average Number of Steps Taken vs 5-min Interval on Weekday/Weekend", y="Average Number of Steps", x="5 minute Interval Times Series")
weekdayinterval
```
